Hazelnut Live: Type-Theoretic Foundations for Running Incomplete Programs

### 2017-06-09 calculus.tex
syntactic consistency
moving around to match use order
adding Delta |- in err and step

\judgbox{\tinconsistent{\htau}{\htau'}}{$\htau$ is inconsistent with $\htau'$}
\begin{mathpar}
    \inferrule[ICBaseArr1]{ }{
      \tinconsistent{\tb}{\tarr{\htau_1}{\htau_2}}
    }

@@

@@

\begin{figure}[!ht]
  \begin{definition}
    $\hasType{\Delta}{\hGamma}{\sigma}{\hGamma'}$ iff for each $\dexp/x \in \sigma$, we have $x : \htau \in \hGamma'$ and $\hasType{\Delta}{\hGamma}{\dexp}{\htau}$.
  \end{definition}
  \caption{substitution type assignment}
\end{figure}

@@

\judgbox{\stepsTo{\dexp_1}{\dexp_2}}{$\dexp_1$ steps to $\dexp_2$}
\begin{mathpar}
\Infer{STHole}
{\stepsTo{\dexp_1}{\dexp_2} }
{\stepsTo{\dhole{\dexp_1}{\mvar}{\subst}}{\dhole{\dexp_2}{\mvar}{\subst}}}
\and
\Infer{STCast}
{\hasType{\Delta}{\emptyset}{\dexp}{\htau_2} \\
 \tconsistent{\tau_1}{\tau_2}}
{\stepsTo{\dcast{\htau_1}{\dexp}}{\dexp}}
\\
\Infer{STAp1}
{\stepsTo{\dexp_1}{\dexp_1'}}
{\stepsTo{\dap{\dexp_1}{\dexp_2}}{\dap{\dexp_1'}{\dexp_2}}}
\and
\Infer{STAp2}
{ \isFinal{\dexp_1} \\ \stepsTo{\dexp_2}{\dexp_2'}}
{\stepsTo{\dap{\dexp_1}{\dexp_2}}{\dap{\dexp_1}{\dexp_2'}}}
\and
\Infer{STAp$\beta$}
{ \isFinal{\dexp_2} }
{\stepsTo{\dapP{\dlam{x}{\htau}{\dexp_1}}{\dexp_2}}{ [\dexp_2/x]\dexp_1 }}
\end{mathpar}
\caption{Structural Dynamics}
\label{fig:isValue}
\label{fig:isIndet}
\label{fig:stepsTo}
\end{figure}

@@

\begin{theorem}[Progress]
If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ then either $\isValue{\dexp}$ or $\isIndet{\dexp}$ or $\isErr{\Delta}{\dexp}$ or $\stepsTo{\dexp}{\dexp'}$.
\end{theorem}

\begin{theorem}[Preservation]
  If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ and $\stepsTo{\dexp}{\dexp'}$ then $\hasType{\Delta}{\emptyset}{\dexp'}{\htau}$. \todo{or maybe up to consistency?}
\end{theorem}

>>>>

\judgbox{\tinconsistent{\htau_1}{\htau_2}}{$\htau_1$ is inconsistent with $\htau_2$}
\begin{mathpar}
    \inferrule[ICBaseArr1]{ }{
      \tinconsistent{\tb}{\tarr{\htau_1}{\htau_2}}
    }

@@

\begin{figure}[!ht]
  \begin{definition}
    $\hasType{\Delta}{\hGamma}{\sigma}{\hGamma'}$ iff for each $\dexp/x \in \sigma$, we have $x : \htau \in \hGamma'$ and $\hasType{\Delta}{\hGamma}{\dexp}{\htau}$.
  \end{definition}
  \caption{substitution type assignment}
  \label{fig:subassign}
\end{figure}

@@

@@

\judgbox{\stepsToD{\Delta}{\dexp_1}{\dexp_2}}{under $\Delta$, $\dexp_1$ steps to $\dexp_2$}
\begin{mathpar}
\Infer{STHole}
{\stepsToD{\Delta}{\dexp_1}{\dexp_2} }
{\stepsToD{\Delta}{\dhole{\dexp_1}{\mvar}{\subst}}{\dhole{\dexp_2}{\mvar}{\subst}}}
\and
\Infer{STCast}
{\hasType{\Delta}{\emptyset}{\dexp}{\htau_2} \\
 \tconsistent{\tau_1}{\tau_2}}
{\stepsToD{\Delta}{\dcast{\htau_1}{\dexp}}{\dexp}}
\\
\Infer{STAp1}
{\stepsToD{\Delta}{\dexp_1}{\dexp_1'}}
{\stepsToD{\Delta}{\dap{\dexp_1}{\dexp_2}}{\dap{\dexp_1'}{\dexp_2}}}
\and
\Infer{STAp2}
{ \isFinal{\dexp_1} \\ \stepsToD{\Delta}{\dexp_2}{\dexp_2'}}
{\stepsToD{\Delta}{\dap{\dexp_1}{\dexp_2}}{\dap{\dexp_1}{\dexp_2'}}}
\and
\Infer{STAp$\beta$}
{ \isFinal{\dexp_2} }
{\stepsToD{\Delta}{\dapP{\dlam{x}{\htau}{\dexp_1}}{\dexp_2}}{ [\dexp_2/x]\dexp_1 }}
\end{mathpar}
\caption{Structural Dynamics}
\label{fig:isValue}
\label{fig:isIndet}
\label{fig:stepsTo}
\end{figure}

@@

\begin{theorem}[Progress]
If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ then either $\isValue{\dexp}$ or $\isIndet{\dexp}$ or $\isErr{\Delta}{\dexp}$ or $\stepsToD{\Delta}{\dexp}{\dexp'}$.
\end{theorem}

\begin{theorem}[Preservation]
  If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ and $\stepsToD{\Delta}{\dexp}{\dexp'}$ then $\hasType{\Delta}{\emptyset}{\dexp'}{\htau}$. \todo{or maybe up to consistency?}
\end{theorem}
